\Title

### 2023-06-14 Grimaldi-etal-BiolCybernetics.tex
Update on Overleaf.

After training our model, we first analyze the weights learned for the different neurons (\seeFig{kernels}). We first notice that these cells are very localized around linear trajectories in the space-delay space, and also show an orientation selectivity perpendicular to motion, similarly to what is found for MT neurons~\citep{deangelis_functional_1999}. Focusing on the positive weights, a strong selectivity along specific motion axes is observed for each of the different kernels. These directions can be easily associated with the direction of motion controlled in the natural images and show a typical anisotropic profile~\citep{kaplan_anisotropic_2013}. The negative weights surround these positive weights, providing an enhanced response. For example, the first kernel shows a strong selectivity for a downward direction of motion. Note that as the delays become larger, two effects can be remarked. First, coefficients become lower which is consistent with the fact that trajectories are defined in a piecewise fashion, such that such decrease provides with an optimal integration with respect to this limited memory. Second, coefficients become less localized compared to the kerenl's spatial profile at short delays, consistent with the average diffusion of information included in the generative model and with the diffusion introduced in motion-based prediction models~\citep{perrinet_motion-based_2012,khoei_flash-lag_2017}. We also observed a strong dependence between the weights reaching the ON polarities and those reaching the OFF polarities. In particular, whenever a weight for a given position and delay is positive for one polarity, it will be negative for the other. This property is due to the way the events are generated and the fact that the luminance cannot increase and decrease at the same time.  Interestingly, the relative organization of the receptive fields in quadrature of phase follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results obtained after showing similar natural scenes with synthetic eye movements~\citep{baudot_animation_2013}.

If one focuses on the interpretation of these kernels in terms of spatiotemporal motifs embedded in the event stream, it can lead to interesting outcomes. In~\citep{grimaldi_robust_2022}, a link between event-based MLR training and Hebbian learning is drawn, allowing to say that the present model will learn its weights according to a presynaptic activity associated to the different motion directions. Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatiotemporal spiking motif. Each voxel in the 3D kernels defines a specific timestamp and a specific address. Consequently, our model is able to detect precise spatiotemporal motifs embedded in the spike train and associated to the different motion directions. The cone shape for the positive weights' distribution highlights a loss of precision for longer delays, i.e. events away in the past. 
%
%%%-----------------------------------------------------------------
%
\subsection{Testing with natural-like textures}
%%%-----------------------------------------------------------------
%%%-----------------------------------------------------------------
%: fig:motion_clouds
%%%-----------------------------------------------------------------
\begin{figure}%[h!]
    \centering
    \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
    \caption{{\bf Role of stimulus parameters in motion detection accuracy.} Accuracy as a function of (from left to right) mean spatial frequency, bandwidth in spatial frequency (from a grating (left) to isotropic textures (right)), bandwidth in orientation (from isotropic textures (left) to grating (right)), bandwidth in speed (from a rigid motion (left) to independent frames (right)). Note that these accuracies are computed both in the case where the orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) and in the generic case where the orientation is independent of direction ('aperture').
    }
    \label{fig:motion_clouds}
\end{figure}
%%%-----------------------------------------------------------------
To further test our model, we will quantify its ability to categorize different motions, i.e., on event streams for which the ground truth motion is known at each instant. Before applying the model to natural images, we will first test the model on simpler, parameterized stimuli. For this purpose, we will use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control, among other parameters, the motion (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}. In particular, we will set the spatial extent and duration of the generated movies similar to the motion task defined above. This procedure defines a set of textures with different spatial properties and different motions chosen from the same set of $XXX8$ directions and $YYY3$ speeds. For each motion, we also varied the texture parameters, such as mean and variance of orientation or spatial frequency content, to provide some naturalistic variability. This method provides a rich dataset of textured movies for which we know the ground truth for the motion.

>>>>

After training our model, let us first analyze the weights learned for the different neurons (\seeFig{kernels}). We first notice by following each spatial pattern from the shortest delay (right) to the longest (left) that the selectivity of these cells are very localized and are transported along linear trajectories in the space-delay space. Focusing on the positive weights, a strong selectivity along specific motion axes is observed for each of the different kernels and these directions can be easily associated with the direction of the corresponding motion. The first kernel, for example, shows a strong selectivity for a downward direction of motion. The opposite weights surround these positive weights, providing an enhanced response. We also observed a strong dependence between the weights reaching the ON polarities and those reaching the OFF polarities. In particular, whenever a weight for a given position and delay is positive for one polarity, it will be negative for the other. This property is due to the way events are generated and the fact that the luminance cannot increase and decrease at the same time.  Interestingly, the relative organization of the receptive fields that we observe is in quadrature of phase and follows a push-pull organization predicted by~\citet{kremkow_push-pull_2016} to explain neurophysiological results obtained after showing similar natural scenes with synthetic eye movements~\citep{baudot_animation_2013}. Finally, we observe that these receptive fields show also a relative selectivity to the orientation perpendicular to motion, similarly to what is found for MT neurons~\citep{deangelis_functional_1999}. This reflects the way events are generated and in particular the so-called aperture problem which implies that an line moving along its axis would generate no change in luminance and therefore generate no event~\citep{perrinet_motion-based_2012}.

If one now widen our focus on the interpretation of these kernels in terms of spatiotemporal motifs embedded in the event stream, these show a prototypical anisotropic profile~\citep{kaplan_anisotropic_2013} adapted to motion detection. In~\citep{grimaldi_robust_2022}, a link was drawn between event-based MLR training and Hebbian learning, allowing to say that the present model will learn its weights according to a presynaptic activity associated to the different motion directions. Each neuron becomes selective to a specific motion direction through the learning of an associated prototypical spatiotemporal spiking motif. Each voxel in the 3D kernels defines a specific property by associating a weight to a position and a delay. Consequently, our model is able to detect precise spatiotemporal motifs embedded in the spike train and associated to the different motion directions. Note that as the delays become larger, two effects can be remarked. First, coefficients become lower which is consistent with the fact that trajectories are defined in a piecewise fashion, such that such decrease provides with an optimal integration with respect to this limited memory. Second, coefficients become less localized compared to the kernel's spatial profile at short delays, consistent with the average diffusion of information included in the generative model and with the diffusion introduced in motion-based prediction models~\citep{perrinet_motion-based_2012,khoei_flash-lag_2017}.  The cone shape for the positive weights' distribution highlights a loss of precision for longer delays, i.e. events away in the past.
%
%%%-----------------------------------------------------------------
%
\subsection{Testing with natural-like textures}
%%%-----------------------------------------------------------------
%%%-----------------------------------------------------------------
%: fig:motion_clouds
%%%-----------------------------------------------------------------
\begin{figure}%[h!]
    \centering
    \includegraphics[width=0.99\linewidth]{figures/motion_clouds.pdf} %
    \caption{{\bf Role of stimulus parameters in motion detection accuracy.} Accuracy as a function of (from left to right) mean spatial frequency, bandwidth in spatial frequency (from a grating (left) to isotropic textures (right)), bandwidth in orientation (from isotropic textures (left) to grating (right)), bandwidth in speed (from a rigid motion (left) to independent frames (right)). Note that these accuracies are computed both in the case where the orientation of the synthetic texture is necessarily perpendicular to the motion ('no aperture' condition) and in the generic case where the orientation is independent of direction ('aperture').
    }
    \label{fig:motion_clouds}
\end{figure}
%%%-----------------------------------------------------------------
To further test our model, we will quantify its ability to categorize different motions, i.e., on event streams for which the ground truth motion is known at each instant. Before applying the model to natural images, we will first test the model on simpler, parameterized stimuli. For this purpose, we will use a set of synthetic visual stimuli, \textit{Motion Clouds}~\citep{leon_motion_2012}, which are natural-like random textures for which we can control, among other parameters, the motion (\seeFig{motion_clouds})~\citep{vacher_bayesian_2018}. In particular, we will set the spatial extent and duration of the generated movies similar to the motion task defined above. This procedure defines a set of textures with different spatial properties and different motions chosen from the same set of $XXX8$ directions and $YYY3$ speeds. For each motion, we also varied the texture parameters, such as mean and variance of orientation or spatial frequency content, to provide some naturalistic variability. This method provides a rich dataset of textured movies for which we know the ground truth for the motion.
